sem

### 2023-01-31 scripts/workshop/sem.Rmd
add text

@@

```{r}
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth() + 
  labs(title = "cc ~ temp")
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth() + 
  labs(title = "log(size) ~ temp")
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth() + 
  labs(title = "Prev ~ temp")
comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
  labs(title = "Prev ~ log(cc)")
comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
  labs(title = "Prev ~ log(size)")
```

# SEMs

In this first iteration, I first propose the relationships independently and then combine them later to generate the SEM. Each relationship is modeled as a linear regression and is evaluated separately. The 'piecewiseSEM' package generates a test statistic (Fisher's C) and and AIC score through a method known as "tests of directed seperation". This method evalutes the relationships between variables that are *un*specified in the model and asks whether these relationships are truly independent or whether some causal relationship may exist. piecewiseSEM relies on a priori knowledge of the system that can be used to rationalize expected relationships in the data. The benefits of piecewiseSEM is that it allows for non-independence of the data and allows for each relationship to be evaluated independently using a more flexible range of modeling options (e.g. GLM and GAM). I develop two sets of SEMs here. One assumes prevalence is only influenced directly by cc and size and the other assumes that water temp may also have a direct effect on prevalence.

@@

```{r}
cc_temp <- lm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)

@@

```{r}
lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
lagged_size_temp <- lm(log10(lag_size) ~ lag_temp + lag_cc, data = clean)
lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
lagged_prev_temp_cc <- lm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)

@@

@@

@@

Lefcheck example
This is example data and code that I got from Lefcheck's tutorial
https://cran.r-project.org/web/packages/piecewiseSEM/vignettes/piecewiseSEM.html#worked-example

@@

Covariance-based SEM

>>>>

# Intro

@@

```{r}
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth() + 
  labs(title = "cc ~ temp")
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth() + 
  labs(title = "log(size) ~ temp")
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth() + 
  labs(title = "Prev ~ temp")
comm_data %>% ggplot(.,aes(x=cc, y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
  labs(title = "Prev ~ log(cc)")
comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
  labs(title = "Prev ~ log(size)")
```

# SEMs intro

In this first iteration, I first propose the relationships independently and then combine them later to generate the SEM. Each relationship is modeled as a linear regression and is evaluated separately. The 'piecewiseSEM' package generates a test statistic (Fisher's C) and and AIC score through a method known as "tests of directed separation". This method evalutes the relationships between variables that are *un*specified in the model and asks whether these relationships are truly independent or whether some causal relationship may exist. piecewiseSEM relies on a priori knowledge of the system that can be used to rationalize expected relationships in the data. The benefits of piecewiseSEM is that it allows for non-independence of the data and allows for each relationship to be evaluated independently using a more flexible range of modeling options (e.g. GLM and GAM). I develop two sets of SEMs here. One assumes prevalence is only influenced directly by cc and size and the other assumes that water temp may also have a direct effect on prevalence.

@@

```{r}
cc_temp <- lm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
prev_temp_cc <- lm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)

@@

```{r}
lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
lagged_size_temp <- lm(log10(lag_size) ~ lag_temp, data = clean)
lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
lagged_prev_temp_cc <- lm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)

@@

# piecewiseSEM

@@

# SEM results and plots

@@

Both SEMs are unable to do the tests of directed separation. I think this may be due to there not being enough variables in the model for this test to work.

# Next Steps

 - Try to incorporate nonlinear regression into model for temperature relationships
  - I want to use a GAM with a smooth term for temperature
 - Check to see if there is a problem without having the directed separation test
  - Still not sure why it says there are no independence claims present when it seems like there is?
  - There seems to be a significant relationship between log10(size) ~ cc that isn't being specified but is showing up under the independence claims result

# Lefcheck example
This is example data and code that I got from Lefcheck's tutorial
https://cran.r-project.org/web/packages/piecewiseSEM/vignettes/piecewiseSEM.html#worked-example

@@

# Covariance-based SEM
